### 2024-12-03 paper/paper.qmd
Added Graphs and Rwender Test

---
title: "Finding the True MVP"
subtitle: "An NBA Analysis Through Linear Regression"
author: 
  - Yan B. Mezhiborsky
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(modelsummary)
library(rstanarm)
library(knitr)
library(bayesplot)
library(StanHeaders)
library(Matrix)
library(lme4)
library(patchwork)
library(kableExtra)

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

# 1986
NBA1986 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1986.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1986[2, ]))
NBA1986 <- NBA1986[-c(1, 2), ]
colnames(NBA1986) <- new_col_names
NBA1986[] <- lapply(NBA1986, type.convert, as.is = TRUE)

# 1987
NBA1987 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1987.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1987[2, ]))
NBA1987 <- NBA1987[-c(1, 2), ]
colnames(NBA1987) <- new_col_names
NBA1987[] <- lapply(NBA1987, type.convert, as.is = TRUE)

# 1988
NBA1988 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1988.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1988[2, ]))
NBA1988 <- NBA1988[-c(1, 2), ]
colnames(NBA1988) <- new_col_names
NBA1988[] <- lapply(NBA1988, type.convert, as.is = TRUE)

# 1989
NBA1989 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1989.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1989[2, ]))
NBA1989 <- NBA1989[-c(1, 2), ]
colnames(NBA1989) <- new_col_names
NBA1989[] <- lapply(NBA1989, type.convert, as.is = TRUE)

# 1990
NBA1990 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1990.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1990[2, ]))
NBA1990 <- NBA1990[-c(1, 2), ]
colnames(NBA1990) <- new_col_names
NBA1990[] <- lapply(NBA1990, type.convert, as.is = TRUE)

# 1991
NBA1991 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1991.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1991[2, ]))
NBA1991 <- NBA1991[-c(1, 2), ]
colnames(NBA1991) <- new_col_names
NBA1991[] <- lapply(NBA1991, type.convert, as.is = TRUE)

# 1992
NBA1992 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1992.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1992[2, ]))
NBA1992 <- NBA1992[-c(1, 2), ]
colnames(NBA1992) <- new_col_names
NBA1992[] <- lapply(NBA1992, type.convert, as.is = TRUE)

# 1993
NBA1993 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1993.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1993[2, ]))
NBA1993 <- NBA1993[-c(1, 2), ]
colnames(NBA1993) <- new_col_names
NBA1993[] <- lapply(NBA1993, type.convert, as.is = TRUE)

# 1994
NBA1994 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1994.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1994[2, ]))
NBA1994 <- NBA1994[-c(1, 2), ]
colnames(NBA1994) <- new_col_names
NBA1994[] <- lapply(NBA1994, type.convert, as.is = TRUE)

# 1995
NBA1995 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1995.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1995[2, ]))
NBA1995 <- NBA1995[-c(1, 2), ]
colnames(NBA1995) <- new_col_names
NBA1995[] <- lapply(NBA1995, type.convert, as.is = TRUE)

# 1996
NBA1996 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1996.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1996[2, ]))
NBA1996 <- NBA1996[-c(1, 2), ]
colnames(NBA1996) <- new_col_names
NBA1996[] <- lapply(NBA1996, type.convert, as.is = TRUE)

# 1997
NBA1997 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1997.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1997[2, ]))
NBA1997 <- NBA1997[-c(1, 2), ]
colnames(NBA1997) <- new_col_names
NBA1997[] <- lapply(NBA1997, type.convert, as.is = TRUE)

# 1998
NBA1998 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1998.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1998[2, ]))
NBA1998 <- NBA1998[-c(1, 2), ]
colnames(NBA1998) <- new_col_names
NBA1998[] <- lapply(NBA1998, type.convert, as.is = TRUE)

# 1999
NBA1999 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1999.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1999[2, ]))
NBA1999 <- NBA1999[-c(1, 2), ]
colnames(NBA1999) <- new_col_names
NBA1999[] <- lapply(NBA1999, type.convert, as.is = TRUE)

# 2000
NBA2000 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2000.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2000[2, ]))
NBA2000 <- NBA2000[-c(1, 2), ]
colnames(NBA2000) <- new_col_names
NBA2000[] <- lapply(NBA2000, type.convert, as.is = TRUE)

# 2001
NBA2001 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2001.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2001[2, ]))
NBA2001 <- NBA2001[-c(1, 2), ]
colnames(NBA2001) <- new_col_names
NBA2001[] <- lapply(NBA2001, type.convert, as.is = TRUE)

# 2002
NBA2002 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2002.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2002[2, ]))
NBA2002 <- NBA2002[-c(1, 2), ]
colnames(NBA2002) <- new_col_names
NBA2002[] <- lapply(NBA2002, type.convert, as.is = TRUE)

# 2003
NBA2003 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2003.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2003[2, ]))
NBA2003 <- NBA2003[-c(1, 2), ]
colnames(NBA2003) <- new_col_names
NBA2003[] <- lapply(NBA2003, type.convert, as.is = TRUE)

# 2004
NBA2004 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2004.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2004[2, ]))
NBA2004 <- NBA2004[-c(1, 2), ]
colnames(NBA2004) <- new_col_names
NBA2004[] <- lapply(NBA2004, type.convert, as.is = TRUE)

# 2005
NBA2005 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2005.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2005[2, ]))
NBA2005 <- NBA2005[-c(1, 2), ]
colnames(NBA2005) <- new_col_names
NBA2005[] <- lapply(NBA2005, type.convert, as.is = TRUE)

# 2006
NBA2006 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2006.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2006[2, ]))
NBA2006 <- NBA2006[-c(1, 2), ]
colnames(NBA2006) <- new_col_names
NBA2006[] <- lapply(NBA2006, type.convert, as.is = TRUE)

# 2007
NBA2007 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2007.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2007[2, ]))
NBA2007 <- NBA2007[-c(1, 2), ]
colnames(NBA2007) <- new_col_names
NBA2007[] <- lapply(NBA2007, type.convert, as.is = TRUE)

# 2008
NBA2008 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2008.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2008[2, ]))
NBA2008 <- NBA2008[-c(1, 2), ]
colnames(NBA2008) <- new_col_names
NBA2008[] <- lapply(NBA2008, type.convert, as.is = TRUE)

# 2009
NBA2009 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2009.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2009[2, ]))
NBA2009 <- NBA2009[-c(1, 2), ]
colnames(NBA2009) <- new_col_names
NBA2009[] <- lapply(NBA2009, type.convert, as.is = TRUE)

# 2010
NBA2010 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2010.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2010[2, ]))
NBA2010 <- NBA2010[-c(1, 2), ]
colnames(NBA2010) <- new_col_names
NBA2010[] <- lapply(NBA2010, type.convert, as.is = TRUE)

# 2011
NBA2011 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2011.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2011[2, ]))
NBA2011 <- NBA2011[-c(1, 2), ]
colnames(NBA2011) <- new_col_names
NBA2011[] <- lapply(NBA2011, type.convert, as.is = TRUE)

# 2012
NBA2012 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2012.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2012[2, ]))
NBA2012 <- NBA2012[-c(1, 2), ]
colnames(NBA2012) <- new_col_names
NBA2012[] <- lapply(NBA2012, type.convert, as.is = TRUE)

# 2013
NBA2013 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2013.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2013[2, ]))
NBA2013 <- NBA2013[-c(1, 2), ]
colnames(NBA2013) <- new_col_names
NBA2013[] <- lapply(NBA2013, type.convert, as.is = TRUE)

# 2014
NBA2014 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2014.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2014[2, ]))
NBA2014 <- NBA2014[-c(1, 2), ]
colnames(NBA2014) <- new_col_names
NBA2014[] <- lapply(NBA2014, type.convert, as.is = TRUE)

# 2015
NBA2015 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2015.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2015[2, ]))
NBA2015 <- NBA2015[-c(1, 2), ]
colnames(NBA2015) <- new_col_names
NBA2015[] <- lapply(NBA2015, type.convert, as.is = TRUE)

# 2016
NBA2016 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2016.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2016[2, ]))
NBA2016 <- NBA2016[-c(1, 2), ]
colnames(NBA2016) <- new_col_names
NBA2016[] <- lapply(NBA2016, type.convert, as.is = TRUE)

# 2017
NBA2017 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2017.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2017[2, ]))
NBA2017 <- NBA2017[-c(1, 2), ]
colnames(NBA2017) <- new_col_names
NBA2017[] <- lapply(NBA2017, type.convert, as.is = TRUE)

# 2018
NBA2018 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2018.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2018[2, ]))
NBA2018 <- NBA2018[-c(1, 2), ]
colnames(NBA2018) <- new_col_names
NBA2018[] <- lapply(NBA2018, type.convert, as.is = TRUE)

# 2019
NBA2019 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2019.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2019[2, ]))
NBA2019 <- NBA2019[-c(1, 2), ]
colnames(NBA2019) <- new_col_names
NBA2019[] <- lapply(NBA2019, type.convert, as.is = TRUE)

# 2020
NBA2020 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2020.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2020[2, ]))
NBA2020 <- NBA2020[-c(1, 2), ]
colnames(NBA2020) <- new_col_names
NBA2020[] <- lapply(NBA2020, type.convert, as.is = TRUE)

# 2021
NBA2021 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2021.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2021[2, ]))
NBA2021 <- NBA2021[-c(1, 2), ]
colnames(NBA2021) <- new_col_names
NBA2021[] <- lapply(NBA2021, type.convert, as.is = TRUE)

# 2022
NBA2022 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2022.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2022[2, ]))
NBA2022 <- NBA2022[-c(1, 2), ]
colnames(NBA2022) <- new_col_names
NBA2022[] <- lapply(NBA2022, type.convert, as.is = TRUE)

# 2023
NBA2023 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2023.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2023[2, ]))
NBA2023 <- NBA2023[-c(1, 2), ]
colnames(NBA2023) <- new_col_names
NBA2023[] <- lapply(NBA2023, type.convert, as.is = TRUE)

# 2024
NBA2024 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2024.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2024[2, ]))
NBA2024 <- NBA2024[-c(1, 2), ]
colnames(NBA2024) <- new_col_names
NBA2024[] <- lapply(NBA2024, type.convert, as.is = TRUE)
```

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

Real_MVP_Data <- read.csv("../data/01-raw_data/NBA MVPs/Real_MVP_Data.csv")

```

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

# Years range
years <- 1986:2024

# Normalize columns and calculate MVP index with WS
for (year in years) {
  dataset_name <- paste0("NBA", year)
  dataset <- get(dataset_name)

  if (all(c("PTS", "AST", "TRB", "STL", "BLK", "WS") %in% colnames(dataset))) {
    averages <- colMeans(dataset[c("PTS", "AST", "TRB", "STL", "BLK", "WS")], na.rm = TRUE)

    dataset$avg_PTS <- averages["PTS"]
    dataset$avg_AST <- averages["AST"]
    dataset$avg_TRB <- averages["TRB"]
    dataset$avg_STL <- averages["STL"]
    dataset$avg_BLK <- averages["BLK"]
    dataset$avg_WS <- averages["WS"]

    dataset$norm_PTS <- dataset$PTS / dataset$avg_PTS
    dataset$norm_AST <- dataset$AST / dataset$avg_AST
    dataset$norm_TRB <- dataset$TRB / dataset$avg_TRB
    dataset$norm_STL <- dataset$STL / dataset$avg_STL
    dataset$norm_BLK <- dataset$BLK / dataset$avg_BLK
    dataset$norm_WS <- dataset$WS / dataset$avg_WS

    dataset$MVP_index <- 0.8 * dataset$norm_PTS +
                         0.25 * dataset$norm_AST +
                         0.25 * dataset$norm_TRB +
                         0.1 * dataset$norm_BLK +
                         0.1 * dataset$norm_STL +
                         0.15 * dataset$norm_WS

    assign(dataset_name, dataset)
  } else {
    stop(paste0("One or more required columns (PTS, AST, TRB, STL, BLK, WS) are missing in dataset for year ", year))
  }
}

# Create a table of the player with the highest MVP index for each year
mvp_table <- data.frame(Year = integer(), Player = character(), MVP_index = numeric(), 
                        PTS = numeric(), AST = numeric(), TRB = numeric(), 
                        BLK = numeric(), STL = numeric(), WS = numeric(), 
                        stringsAsFactors = FALSE)

# Iterate over each year to find the top player
for (year in years) {
  dataset_name <- paste0("NBA", year)
  dataset <- get(dataset_name)

  if ("MVP_index" %in% colnames(dataset)) {
    max_row <- dataset[which.max(dataset$MVP_index), ]

    mvp_entry <- data.frame(
      Year = year, 
      Player = max_row$Player, 
      MVP_index = max_row$MVP_index,
      PTS = max_row$PTS,
      AST = max_row$AST,
      TRB = max_row$TRB,
      BLK = max_row$BLK,
      STL = max_row$STL,
      WS = max_row$WS,
      stringsAsFactors = FALSE
    )

    mvp_table <- rbind(mvp_table, mvp_entry)
  } else {
    stop(paste0("MVP_index column is missing in dataset for year ", year))
  }
}

nba_list <- list()

for (year in years) {
  dataset_name <- paste0("NBA", year)
  dataset <- get(dataset_name)
  dataset$Year <- year
  nba_list[[as.character(year)]] <- dataset
}

nba_master <- do.call(rbind, nba_list)
write.csv(nba_master, "../data/02-analysis_data/nba_master.csv", row.names = FALSE)

merged_real_mvp <- merge(
  Real_MVP_Data, 
  nba_master[, c("Player", "Age", "MVP_index")],  # Select only relevant columns from nba_master
  by = c("Player", "Age"), 
  all.x = TRUE
)

merged_real_mvp <- merged_real_mvp %>%
  mutate(
    Year = case_when(
      # If entry matches "YYYY-YY", extract the first 4 characters as the year
      grepl("^\\d{4}-\\d{2}$", Season) ~ as.numeric(substr(Season, 1, 4)) + 1,

      # If entry matches "Mmm-YY", extract the last 2 characters, convert to 20YY, and add 1
      grepl("^[a-zA-Z]{3}-\\d{2}$", Season) ~ as.numeric(paste0("20", substr(Season, 5, 6))) + 1,

      # Default: NA for non-matching formats
      TRUE ~ NA_real_
    )
  )

# Convert Year column to numeric
merged_real_mvp$Year <- as.numeric(merged_real_mvp$Year)

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \pi_i + \alpha_i + \rho_i + \beta_i + \xi_i + \omega_i \\
\mu &\sim \mbox{Normal}(0, 2.5) \\
\pi &\sim \mbox{Normal}(0, 2.5) \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\rho &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\xi &\sim \mbox{Normal}(0, 2.5) \\
\omega &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

```{r}
# Fit a linear regression model
lm_model <- lm(MVP_index ~ PTS + AST + TRB + BLK + STL + WS, data = nba_master)

# View the summary of the model
summary(lm_model)

next_year_mvp_stats <- mvp_table %>%
  filter(Year == 2024) %>%  # Replace with the target year
  select(PTS, AST, TRB, BLK, STL, WS)

# Predict MVP_index for the next year's MVP
predicted_index_mvp_table <- predict(lm_model, newdata = next_year_mvp_stats)
print(paste("Predicted MVP Index (mvp_table):", predicted_index_mvp_table))

next_year_real_mvp_stats <- merged_real_mvp %>%
  filter(Year == 2024) %>%  # Replace with the target year
  select(PTS, AST, TRB, BLK, STL, WS)

# Predict MVP_index for the next year's MVP
predicted_index_real_mvp <- predict(lm_model, newdata = next_year_real_mvp_stats)
print(paste("Predicted MVP Index (merged_real_mvp):", predicted_index_real_mvp))

```

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

### Model Results
```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false

# Filter MVPs from 2014 to 2024 for both datasets
mvp_table_filtered <- mvp_table %>%
  filter(Year >= 2014 & Year <= 2024) %>%
  select(Year, Player, MVP_index)

real_mvp_filtered <- merged_real_mvp %>%
  filter(Year >= 2014 & Year <= 2024) %>%
  select(Year, Player, MVP_index)

# Predict MVP index for 2025 and add it to both tables
next_year_mvp_stats <- mvp_table %>%
  filter(Year == 2024) %>%  # Use 2024 stats as proxy
  select(PTS, AST, TRB, BLK, STL, WS)

predicted_index_mvp_table <- NA
if (nrow(next_year_mvp_stats) > 0) {
  predicted_index_mvp_table <- predict(lm_model, newdata = next_year_mvp_stats)
}

mvp_table_filtered <- bind_rows(
  mvp_table_filtered,
  data.frame(
    Year = 2025,
    Player = "Predicted MVP",
    MVP_index = predicted_index_mvp_table
  )
)

next_year_real_mvp_stats <- merged_real_mvp %>%
  filter(Year == 2024) %>%  # Use 2024 stats as proxy
  select(PTS, AST, TRB, BLK, STL, WS)

predicted_index_real_mvp <- NA
if (nrow(next_year_real_mvp_stats) > 0) {
  predicted_index_real_mvp <- predict(lm_model, newdata = next_year_real_mvp_stats)
}

real_mvp_filtered <- bind_rows(
  real_mvp_filtered,
  data.frame(
    Year = 2025,
    Player = "Predicted MVP",
    MVP_index = predicted_index_real_mvp
  )
)

# Rename columns for clarity
mvp_table_filtered <- mvp_table_filtered %>%
  rename(
    `Simulated Data - Player` = Player,
    `Simulated Data - MVP Index` = MVP_index
  )

real_mvp_filtered <- real_mvp_filtered %>%
  rename(
    `Historical Data - Player` = Player,
    `Historical Data - MVP Index` = MVP_index
  )

# Join the two tables by Year
combined_table <- full_join(
  mvp_table_filtered,
  real_mvp_filtered,
  by = "Year"
)

# Create a styled table using kable
combined_table %>%
  kable(
    format = "html",
    caption = "Combined MVP Data (2014-2025)",
    col.names = c(
      "Year",
      "Simulated Data - Player",
      "Simulated Data - MVP Index",
      "Historical Data - Player",
      "Historical Data - MVP Index"
    )
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  add_header_above(c(" " = 1, "Simulated Data" = 2, "Historical Data" = 2))

```

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

ggplot(merged_real_mvp, aes(x = Year, y = MVP_index)) +
  geom_point(color = "blue", size = 3, alpha = 0.7) +
  labs(
    title = "Scatter: Year vs. MVP Index (Merged Real MVPs)",
    x = "Year",
    y = "MVP Index"
  ) +
  coord_cartesian(ylim = c(1.7, 2.8)) +  # Maintain the same y-axis range
  theme_minimal()

ggplot(mvp_table, aes(x = Year, y = MVP_index)) +
  geom_point(color = "red", size = 3, alpha = 0.7) +
  labs(
    title = "Scatter: Year vs. MVP Index (MVP Table)",
    x = "Year",
    y = "MVP Index"
  ) +
  coord_cartesian(ylim = c(1.7, 2.8)) +  # Maintain the same y-axis range
  theme_minimal()

```

```{r}

# Add a source column to the datasets
mvp_table <- mvp_table %>%
  mutate(Source = "Simulated MVPs")

merged_real_mvp <- merged_real_mvp %>%
  mutate(Source = "Historical MVPs")

# Combine datasets
combined_data <- bind_rows(
  mvp_table %>% select(Year, PTS, TRB, AST, Source),
  merged_real_mvp %>% select(Year, PTS, TRB, AST, Source)
)

# Plot for PTS
pts_plot <- ggplot(combined_data, aes(x = Year, y = PTS, group = Source, color = Source)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ Source) +
  labs(
    title = "Year vs. PTS",
    x = "Year",
    y = "PTS"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Plot for TRB
trb_plot <- ggplot(combined_data, aes(x = Year, y = TRB, group = Source, color = Source)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ Source) +
  labs(
    title = "Year vs. TRB",
    x = "Year",
    y = "TRB"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Plot for AST
ast_plot <- ggplot(combined_data, aes(x = Year, y = AST, group = Source, color = Source)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ Source) +
  labs(
    title = "Year vs. AST",
    x = "Year",
    y = "AST"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Display plots separately
print(pts_plot)
print(trb_plot)
print(ast_plot)

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

```

\newpage

# References