### 2024-12-04 paper/paper.qmd
Final Edits

---
title: "Finding the True MVP"
subtitle: "An NBA Analysis Through Linear Regression"
author: 
  - Yan Mezhiborsky
thanks: "Code and data are available at: [https://github.com/Mezhi18/NBAMVP.]"
date: today
date-format: long
abstract: "This paper examines the methodology behind the NBA’s Most Valuable Player (MVP) award, introducing the MVP Index as a data-driven measure to evaluate player performance. The MVP Index combines key statistics, including points, assists, rebounds, and win shares, to provide a comprehensive metric of value. Using historical data and statistical modeling, we identify discrepancies between actual MVP winners and those with the highest MVP Index, highlighting the influence of narrative and qualitative factors on voting outcomes. The findings underscore the need for greater transparency and consistency in the MVP selection process while offering a framework for more objective evaluations."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

library(tidyverse)
library(rstanarm)
library(dplyr)
library(janitor)
library(lubridate)
library(modelsummary)
library(knitr)
library(StanHeaders)
library(kableExtra)

```

# Introduction {#sec-intro}

Basketball’s Most Valuable Player (MVP) award represents one of the most prestigious accolades in professional sports, honoring the individual deemed to have contributed the most to their team’s success over a season. While the selection process has traditionally relied on expert voting and narrative-driven considerations, the question of whether the MVP truly reflects a player's value remains open to debate. This paper seeks to address this issue by evaluating the award through a data-driven lens, introducing the MVP Index as a novel measure of player impact. By exploring historical trends, inconsistencies, and predictive modeling, this research aims to shed light on how MVP selections align—or fail to align—with statistical measures of excellence.

At the heart of this analysis is the estimand, the MVP Index, which combines key player performance metrics into a single, normalized statistic. This index incorporates points, assists, rebounds, blocks, steals, and win shares, weighted to reflect their relative importance. The MVP Index provides a quantitative foundation for evaluating player performance, offering an alternative to the sometimes subjective criteria that influence MVP voting. The study employs statistical modeling, with linear regression used to predict the MVP Index for future seasons, enabling a comparison between the players with the best statistical performance and those who ultimately win the award.

The results of the analysis reveal several significant insights. In some seasons, the MVP Index aligns closely with the actual MVP winner, validating the selection process. However, there are also notable discrepancies where players with superior statistical performances were overlooked in favor of candidates with compelling narratives or team success factors. These findings suggest that while data can provide clarity, the MVP award is often shaped by qualitative elements that extend beyond pure statistics.

Understanding these patterns is not only critical for fans and analysts but also holds implications for broader debates about transparency and fairness in sports. The MVP award significantly impacts player legacies, influencing Hall of Fame considerations, sponsorship deals, and even contract negotiations. A more robust and transparent framework for assessing MVP candidates could enhance the credibility of the award and ensure that it reflects the contributions of players who truly make a difference.

The remainder of this paper is structured as follows. Section @sec-data outlines the data sources, variables, and processing steps used to construct the MVP Index. Section @sec-model details our model, explaining its development and how it is applied to predict future statistics. In Section @sec-results, we analyze MVP statistics over nearly 40 years, examining how they have evolved and highlighting important discrepancies between our model and historical MVP selections. Finally, Section @sec-discussion discusses the broader implications of these findings and acknowledges the limitations of this study.

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR] along with several packages to clean and manipulate our data, including [@tidyverse], [@rstanarm], [@dplyr], [@janitor], [@lubridate], [@modelsummary], [@knitr], [@StanHeaders], and [@kableExtra]. These tools allowed us to clean the data, add years to datasets that only included basketball seasons, and generate well-formatted tables that go beyond the basic functionality of traditional R software.

Our data was collected from [@bbref], focusing on every player who received a vote for any All-NBA Team since 1986. This includes votes for the First, Second, and Third All-NBA Teams. We chose this approach to ensure a larger and more comprehensive sample, as limiting the scope to players who received MVP votes would have resulted in a smaller dataset. Additionally, we created a new statistic called MVP_Index, which is defined in @sec-variables. This statistic was used to evaluate players and compare them to the official NBA MVPs since 1986, providing insights into alternative MVP selections.

Building on the methodologies outlined in [@tellingstories], we employed various techniques to analyze the data. These included statistical models, visualizations, tables, and other methods of data analysis to identify players who could have been MVPs and to predict what future MVPs might look like, based on their statistical profiles.

## Variables {#sec-variables}

### All NBA Variables

As we collected data from each year since 1986 individually, we have utilized the following variables:

-   `Team:` Whether the player was in the 1st, 2nd, or 3rd All NBA Teams or OVR if they did not make the top three teams but still received a vote.
-   `Pos:` The player's position including the generalized positions G, F, C, for Guard, Forward, and Center respectively.
-   `Player:` The name of the NBA player.
-   `Age:` The age of the NBA player.
-   `Tm:` The three letter designation for the NBA team the player played for the year.
-   `G:` Games played by the player.
-   `MP:` Minutes played per game.
-   `PTS:` Points per game.
-   `TRB:` Total Rebounds per game, including offensive and defensive rebounds.
-   `BLK:` Blocks per game
-   `STL:` Steals Per game
-   `WS:` The amount of wins allocated to the individual player

As each year has an individual data set we took PTS, TRB, AST, BLK, STL,a nd WS took the average for that year and normalized each player's statistics by dividing them by the average of that year with this we have created the MVP Index which we have defined as:

$$MVP_{Index} = 0.8PTS_{Norm} + 0.25AST_{Norm} + 0.25TRB_{Norm} + 0.1STL_{Norm} + 0.1BLK_{Norm} + 0.15WS_{Norm}$$

With this `MVP Index` we have created our own list of who could have been MVP base on statistics for that year, essentially comparing how good the player was to othe rplayers in the basketball season.

### Historical MVP Variable

For our second set of date we have all MVPs since 1986 also sourced from [@bbref], and the variables that have been used are the following:

-   `Player:` The name of the NBA player.
-   `Year:` The year in which the player was selected as MVP.
-   `Tm:` The three letter designation for the NBA team the player played for the year.
-   `G:` Games played by the player.
-   `MP:` Minutes played per game.
-   `PTS:` Points per game.
-   `TRB:` Total Rebounds per game, including offensive and defensive rebounds.
-   `BLK:` Blocks per game
-   `STL:` Steals Per game
-   `WS:` The amount of wins allocated to the individual player

## Measurement

Data collection in sports, and basketball in particular, is both straightforward and meticulous. Every NBA game has dedicated statisticians whose sole responsibility is to record player statistics in real-time. Whether it is a made shot, an assist, a rebound, a steal, or a block, every action is logged systematically to ensure accuracy and completeness. This process creates a detailed statistical profile for each player, capturing their performance throughout the game.

Over the course of a standard 82-game season, these statistics accumulate into a comprehensive dataset. However, the number of games played can vary due to external factors such as pandemics like COVID-19, injuries that sideline players, or trades that impact availability. Despite these variations, the collected statistics remain consistent in format and provide a foundation for deeper analysis.

## Outcome variable: MVP Index

Our outcome variable is the same as our estimand in @sec-intro. The MVP Index is dependent on our other statistics: Points, Assists, Rebounds, Steals, Blocks, and Win Shares. As we are studying MVPs in our paper, this is our variable of interest as it represents how good a player was in comparison to the other best players in the NBA that year, from the sample of players who received any All-NBA Team vote.

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

# 1986
NBA1986 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1986.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1986[2, ]))
NBA1986 <- NBA1986[-c(1, 2), ]
colnames(NBA1986) <- new_col_names
NBA1986[] <- lapply(NBA1986, type.convert, as.is = TRUE)

# 1987
NBA1987 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1987.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1987[2, ]))
NBA1987 <- NBA1987[-c(1, 2), ]
colnames(NBA1987) <- new_col_names
NBA1987[] <- lapply(NBA1987, type.convert, as.is = TRUE)

# 1988
NBA1988 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1988.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1988[2, ]))
NBA1988 <- NBA1988[-c(1, 2), ]
colnames(NBA1988) <- new_col_names
NBA1988[] <- lapply(NBA1988, type.convert, as.is = TRUE)

# 1989
NBA1989 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1989.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1989[2, ]))
NBA1989 <- NBA1989[-c(1, 2), ]
colnames(NBA1989) <- new_col_names
NBA1989[] <- lapply(NBA1989, type.convert, as.is = TRUE)

# 1990
NBA1990 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1990.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1990[2, ]))
NBA1990 <- NBA1990[-c(1, 2), ]
colnames(NBA1990) <- new_col_names
NBA1990[] <- lapply(NBA1990, type.convert, as.is = TRUE)

# 1991
NBA1991 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1991.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1991[2, ]))
NBA1991 <- NBA1991[-c(1, 2), ]
colnames(NBA1991) <- new_col_names
NBA1991[] <- lapply(NBA1991, type.convert, as.is = TRUE)

# 1992
NBA1992 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1992.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1992[2, ]))
NBA1992 <- NBA1992[-c(1, 2), ]
colnames(NBA1992) <- new_col_names
NBA1992[] <- lapply(NBA1992, type.convert, as.is = TRUE)

# 1993
NBA1993 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1993.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1993[2, ]))
NBA1993 <- NBA1993[-c(1, 2), ]
colnames(NBA1993) <- new_col_names
NBA1993[] <- lapply(NBA1993, type.convert, as.is = TRUE)

# 1994
NBA1994 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1994.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1994[2, ]))
NBA1994 <- NBA1994[-c(1, 2), ]
colnames(NBA1994) <- new_col_names
NBA1994[] <- lapply(NBA1994, type.convert, as.is = TRUE)

# 1995
NBA1995 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1995.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1995[2, ]))
NBA1995 <- NBA1995[-c(1, 2), ]
colnames(NBA1995) <- new_col_names
NBA1995[] <- lapply(NBA1995, type.convert, as.is = TRUE)

# 1996
NBA1996 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1996.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1996[2, ]))
NBA1996 <- NBA1996[-c(1, 2), ]
colnames(NBA1996) <- new_col_names
NBA1996[] <- lapply(NBA1996, type.convert, as.is = TRUE)

# 1997
NBA1997 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1997.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1997[2, ]))
NBA1997 <- NBA1997[-c(1, 2), ]
colnames(NBA1997) <- new_col_names
NBA1997[] <- lapply(NBA1997, type.convert, as.is = TRUE)

# 1998
NBA1998 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1998.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1998[2, ]))
NBA1998 <- NBA1998[-c(1, 2), ]
colnames(NBA1998) <- new_col_names
NBA1998[] <- lapply(NBA1998, type.convert, as.is = TRUE)

# 1999
NBA1999 <- read.csv("../data/01-raw_data/All NBA/AllNBA_1999.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA1999[2, ]))
NBA1999 <- NBA1999[-c(1, 2), ]
colnames(NBA1999) <- new_col_names
NBA1999[] <- lapply(NBA1999, type.convert, as.is = TRUE)

# 2000
NBA2000 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2000.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2000[2, ]))
NBA2000 <- NBA2000[-c(1, 2), ]
colnames(NBA2000) <- new_col_names
NBA2000[] <- lapply(NBA2000, type.convert, as.is = TRUE)

# 2001
NBA2001 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2001.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2001[2, ]))
NBA2001 <- NBA2001[-c(1, 2), ]
colnames(NBA2001) <- new_col_names
NBA2001[] <- lapply(NBA2001, type.convert, as.is = TRUE)

# 2002
NBA2002 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2002.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2002[2, ]))
NBA2002 <- NBA2002[-c(1, 2), ]
colnames(NBA2002) <- new_col_names
NBA2002[] <- lapply(NBA2002, type.convert, as.is = TRUE)

# 2003
NBA2003 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2003.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2003[2, ]))
NBA2003 <- NBA2003[-c(1, 2), ]
colnames(NBA2003) <- new_col_names
NBA2003[] <- lapply(NBA2003, type.convert, as.is = TRUE)

# 2004
NBA2004 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2004.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2004[2, ]))
NBA2004 <- NBA2004[-c(1, 2), ]
colnames(NBA2004) <- new_col_names
NBA2004[] <- lapply(NBA2004, type.convert, as.is = TRUE)

# 2005
NBA2005 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2005.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2005[2, ]))
NBA2005 <- NBA2005[-c(1, 2), ]
colnames(NBA2005) <- new_col_names
NBA2005[] <- lapply(NBA2005, type.convert, as.is = TRUE)

# 2006
NBA2006 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2006.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2006[2, ]))
NBA2006 <- NBA2006[-c(1, 2), ]
colnames(NBA2006) <- new_col_names
NBA2006[] <- lapply(NBA2006, type.convert, as.is = TRUE)

# 2007
NBA2007 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2007.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2007[2, ]))
NBA2007 <- NBA2007[-c(1, 2), ]
colnames(NBA2007) <- new_col_names
NBA2007[] <- lapply(NBA2007, type.convert, as.is = TRUE)

# 2008
NBA2008 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2008.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2008[2, ]))
NBA2008 <- NBA2008[-c(1, 2), ]
colnames(NBA2008) <- new_col_names
NBA2008[] <- lapply(NBA2008, type.convert, as.is = TRUE)

# 2009
NBA2009 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2009.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2009[2, ]))
NBA2009 <- NBA2009[-c(1, 2), ]
colnames(NBA2009) <- new_col_names
NBA2009[] <- lapply(NBA2009, type.convert, as.is = TRUE)

# 2010
NBA2010 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2010.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2010[2, ]))
NBA2010 <- NBA2010[-c(1, 2), ]
colnames(NBA2010) <- new_col_names
NBA2010[] <- lapply(NBA2010, type.convert, as.is = TRUE)

# 2011
NBA2011 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2011.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2011[2, ]))
NBA2011 <- NBA2011[-c(1, 2), ]
colnames(NBA2011) <- new_col_names
NBA2011[] <- lapply(NBA2011, type.convert, as.is = TRUE)

# 2012
NBA2012 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2012.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2012[2, ]))
NBA2012 <- NBA2012[-c(1, 2), ]
colnames(NBA2012) <- new_col_names
NBA2012[] <- lapply(NBA2012, type.convert, as.is = TRUE)

# 2013
NBA2013 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2013.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2013[2, ]))
NBA2013 <- NBA2013[-c(1, 2), ]
colnames(NBA2013) <- new_col_names
NBA2013[] <- lapply(NBA2013, type.convert, as.is = TRUE)

# 2014
NBA2014 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2014.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2014[2, ]))
NBA2014 <- NBA2014[-c(1, 2), ]
colnames(NBA2014) <- new_col_names
NBA2014[] <- lapply(NBA2014, type.convert, as.is = TRUE)

# 2015
NBA2015 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2015.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2015[2, ]))
NBA2015 <- NBA2015[-c(1, 2), ]
colnames(NBA2015) <- new_col_names
NBA2015[] <- lapply(NBA2015, type.convert, as.is = TRUE)

# 2016
NBA2016 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2016.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2016[2, ]))
NBA2016 <- NBA2016[-c(1, 2), ]
colnames(NBA2016) <- new_col_names
NBA2016[] <- lapply(NBA2016, type.convert, as.is = TRUE)

# 2017
NBA2017 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2017.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2017[2, ]))
NBA2017 <- NBA2017[-c(1, 2), ]
colnames(NBA2017) <- new_col_names
NBA2017[] <- lapply(NBA2017, type.convert, as.is = TRUE)

# 2018
NBA2018 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2018.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2018[2, ]))
NBA2018 <- NBA2018[-c(1, 2), ]
colnames(NBA2018) <- new_col_names
NBA2018[] <- lapply(NBA2018, type.convert, as.is = TRUE)

# 2019
NBA2019 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2019.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2019[2, ]))
NBA2019 <- NBA2019[-c(1, 2), ]
colnames(NBA2019) <- new_col_names
NBA2019[] <- lapply(NBA2019, type.convert, as.is = TRUE)

# 2020
NBA2020 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2020.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2020[2, ]))
NBA2020 <- NBA2020[-c(1, 2), ]
colnames(NBA2020) <- new_col_names
NBA2020[] <- lapply(NBA2020, type.convert, as.is = TRUE)

# 2021
NBA2021 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2021.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2021[2, ]))
NBA2021 <- NBA2021[-c(1, 2), ]
colnames(NBA2021) <- new_col_names
NBA2021[] <- lapply(NBA2021, type.convert, as.is = TRUE)

# 2022
NBA2022 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2022.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2022[2, ]))
NBA2022 <- NBA2022[-c(1, 2), ]
colnames(NBA2022) <- new_col_names
NBA2022[] <- lapply(NBA2022, type.convert, as.is = TRUE)

# 2023
NBA2023 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2023.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2023[2, ]))
NBA2023 <- NBA2023[-c(1, 2), ]
colnames(NBA2023) <- new_col_names
NBA2023[] <- lapply(NBA2023, type.convert, as.is = TRUE)

# 2024
NBA2024 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2024.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2024[2, ]))
NBA2024 <- NBA2024[-c(1, 2), ]
colnames(NBA2024) <- new_col_names
NBA2024[] <- lapply(NBA2024, type.convert, as.is = TRUE)
```

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

Real_MVP_Data <- read.csv("../data/01-raw_data/NBA MVPs/Real_MVP_Data.csv")

```

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

# Years range
years <- 1986:2024

# Normalize columns and calculate MVP index with WS
for (year in years) {
  dataset_name <- paste0("NBA", year)
  dataset <- get(dataset_name)

  if (all(c("PTS", "AST", "TRB", "STL", "BLK", "WS") %in% colnames(dataset))) {
    averages <- colMeans(dataset[c("PTS", "AST", "TRB", "STL", "BLK", "WS")], na.rm = TRUE)

    dataset$avg_PTS <- averages["PTS"]
    dataset$avg_AST <- averages["AST"]
    dataset$avg_TRB <- averages["TRB"]
    dataset$avg_STL <- averages["STL"]
    dataset$avg_BLK <- averages["BLK"]
    dataset$avg_WS <- averages["WS"]

    dataset$norm_PTS <- dataset$PTS / dataset$avg_PTS
    dataset$norm_AST <- dataset$AST / dataset$avg_AST
    dataset$norm_TRB <- dataset$TRB / dataset$avg_TRB
    dataset$norm_STL <- dataset$STL / dataset$avg_STL
    dataset$norm_BLK <- dataset$BLK / dataset$avg_BLK
    dataset$norm_WS <- dataset$WS / dataset$avg_WS

    dataset$MVP_index <- 0.8 * dataset$norm_PTS +
                         0.25 * dataset$norm_AST +
                         0.25 * dataset$norm_TRB +
                         0.1 * dataset$norm_BLK +
                         0.1 * dataset$norm_STL +
                         0.15 * dataset$norm_WS

    assign(dataset_name, dataset)
  } else {
    stop(paste0("One or more required columns (PTS, AST, TRB, STL, BLK, WS) are missing in dataset for year ", year))
  }
}

# Create a table of the player with the highest MVP index for each year
mvp_table <- data.frame(Year = integer(), Player = character(), MVP_index = numeric(), 
                        PTS = numeric(), AST = numeric(), TRB = numeric(), 
                        BLK = numeric(), STL = numeric(), WS = numeric(), 
                        stringsAsFactors = FALSE)

# Iterate over each year to find the top player
for (year in years) {
  dataset_name <- paste0("NBA", year)
  dataset <- get(dataset_name)

  if ("MVP_index" %in% colnames(dataset)) {
    max_row <- dataset[which.max(dataset$MVP_index), ]

    mvp_entry <- data.frame(
      Year = year, 
      Player = max_row$Player, 
      MVP_index = max_row$MVP_index,
      PTS = max_row$PTS,
      AST = max_row$AST,
      TRB = max_row$TRB,
      BLK = max_row$BLK,
      STL = max_row$STL,
      WS = max_row$WS,
      stringsAsFactors = FALSE
    )

    mvp_table <- rbind(mvp_table, mvp_entry)
  } else {
    stop(paste0("MVP_index column is missing in dataset for year ", year))
  }
}

write.csv(mvp_table, "../data/02-analysis_data/mvp_table.csv", row.names = FALSE)

nba_list <- list()

for (year in years) {
  dataset_name <- paste0("NBA", year)
  dataset <- get(dataset_name)
  dataset$Year <- year
  nba_list[[as.character(year)]] <- dataset
}

nba_master <- do.call(rbind, nba_list)
write.csv(nba_master, "../data/02-analysis_data/nba_master.csv", row.names = FALSE)

merged_real_mvp <- merge(
  Real_MVP_Data, 
  nba_master[, c("Player", "Age", "MVP_index")],  # Select only relevant columns from nba_master
  by = c("Player", "Age"), 
  all.x = TRUE
)

merged_real_mvp <- merged_real_mvp %>%
  mutate(
    Year = case_when(
      grepl("^\\d{4}-\\d{2}$", Season) ~ as.numeric(substr(Season, 1, 4)) + 1,
      grepl("^[a-zA-Z]{3}-\\d{2}$", Season) ~ as.numeric(paste0("20", substr(Season, 5, 6))) + 1,
      TRUE ~ NA_real_
    )
  )

# Convert Year column to numeric
merged_real_mvp$Year <- as.numeric(merged_real_mvp$Year)

```

# Model {#sec-model}

The goal of our modeling strategy is twofold. Firstly, we will use a linear regression model to predict the MVP Index of the MVP in the following year, regardless of the player. This will allow us to estimate how much better the 2025 MVP is likely to perform compared to his fellow NBA players. Secondly, we will utilize another dataset that we created by identifying players with the highest MVP Index in their respective years, whether they won the MVP award or not. By combining these predictions, we can assess whether the player with the highest MVP Index is likely to become the next NBA MVP.

Here, we briefly describe the linear analysis model used to predict the MVP Index of the next NBA MVP. For our model, we rely on the collected data described in @sec-data.

Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the MVP Index. Then $\pi_i$ as the points per game, then $\alpha_i$ as assists per game, next we have $\rho_i$ as rebounds per game. Furthermore, we have $\beta_i$ as blocks per game, then we have $\xi_i$ for steals per game, lastly we have $\omega_i$ for win shares. With these variables we will be predicting the MVP Index for the following NBA season.

\begin{align}
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \pi_i + \alpha_i + \rho_i + \beta_i + \xi_i + \omega_i \\
\pi &\sim \mbox{Normal}(0, 2.5) \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\rho &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\xi &\sim \mbox{Normal}(0, 2.5) \\
\omega &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

# create a linear regression model
lm_model <- lm(MVP_index ~ PTS + AST + TRB + BLK + STL + WS, data = nba_master)

# View the summary of the model
summary(lm_model)

next_year_mvp_stats <- mvp_table %>%
  filter(Year == 2024) %>%  # Replace with the target year
  select(PTS, AST, TRB, BLK, STL, WS)

# Predict MVP_index for the next year's MVP
predicted_index_mvp_table <- predict(lm_model, newdata = next_year_mvp_stats)
print(paste("Predicted MVP Index (mvp_table):", predicted_index_mvp_table))

next_year_real_mvp_stats <- merged_real_mvp %>%
  filter(Year == 2024) %>%  # Replace with the target year
  select(PTS, AST, TRB, BLK, STL, WS)

# Predict MVP_index for the next year's MVP
predicted_index_real_mvp <- predict(lm_model, newdata = next_year_real_mvp_stats)

saveRDS(lm_model, "../models/lm_model.rds")
```

### Model justification

We expect a positive relationship between MVP Index and points, assists, rebounds, blocks, steals, and win shares, as these variables collectively define the MVP Index. Additionally, we anticipate a positive relationship because all these statistics indicate strong player performance. The larger any one of these statistics is, the larger the MVP Index will be for the respective player. However, a player who does not acquire many blocks or assists will not necessarily have a low MVP rating; they will simply need to compensate with higher performance in other areas. As shown in @sec-model-details, the model has a p-value of less than $2.2 \times 10^{-16}$, which is very low and gives us confidence that the model is accurate.

### Model Results

Below, we reference @tbl-modelresults, which presents the MVPs from the last 10 years, along with an additional unknown MVP with a projected MVP Index score. This table includes both historical data with past MVP winners and data we have created independently. Our data identifies players with the highest MVP Index within their respective years, which may not necessarily be the same players who were officially awarded the MVP that year.

As seen in the table, the historical data predicts a player to have an MVP Index of 2.40, while our data suggests that the player with the highest MVP Index will have a score of 2.45. This indicates it is likely that the player with the highest MVP Index will not be the one selected as the MVP. For example, in the 2024 season, the MVP was awarded to Nikola Jokic, who had an MVP Index score of 2.16. However, Luka Doncic, another strong contender for the award, had an MVP Index of 2.20, which was higher than Jokic's score, further highlighting this discrepancy.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-modelresults
#| fig-cap: "Historical MVP and Player with highest MVP Index since 2014"

# Filter MVPs from 2014 to 2024 for both datasets
mvp_table_filtered <- mvp_table %>%
  filter(Year >= 2014 & Year <= 2024) %>%
  select(Year, Player, MVP_index)

real_mvp_filtered <- merged_real_mvp %>%
  filter(Year >= 2014 & Year <= 2024) %>%
  select(Year, Player, MVP_index)

# Predict MVP index for 2025 and add it to both tables
next_year_mvp_stats <- mvp_table %>%
  filter(Year == 2024) %>%  # Use 2024 stats as proxy
  select(PTS, AST, TRB, BLK, STL, WS)

predicted_index_mvp_table <- NA
if (nrow(next_year_mvp_stats) > 0) {
  predicted_index_mvp_table <- predict(lm_model, newdata = next_year_mvp_stats)
}

mvp_table_filtered <- bind_rows(
  mvp_table_filtered,
  data.frame(
    Year = 2025,
    Player = "Predicted MVP",
    MVP_index = predicted_index_mvp_table
  )
)

next_year_real_mvp_stats <- merged_real_mvp %>%
  filter(Year == 2024) %>%  # Use 2024 stats as proxy
  select(PTS, AST, TRB, BLK, STL, WS)

predicted_index_real_mvp <- NA
if (nrow(next_year_real_mvp_stats) > 0) {
  predicted_index_real_mvp <- predict(lm_model, newdata = next_year_real_mvp_stats)
}

real_mvp_filtered <- bind_rows(
  real_mvp_filtered,
  data.frame(
    Year = 2025,
    Player = "Predicted MVP",
    MVP_index = predicted_index_real_mvp
  )
)

# Rename columns for clarity
mvp_table_filtered <- mvp_table_filtered %>%
  rename(
    `Simulated Data - Player` = Player,
    `Simulated Data - MVP Index` = MVP_index
  )

real_mvp_filtered <- real_mvp_filtered %>%
  rename(
    `Historical Data - Player` = Player,
    `Historical Data - MVP Index` = MVP_index
  )

# Join the two tables by Year
combined_table <- full_join(
  mvp_table_filtered,
  real_mvp_filtered,
  by = "Year"
)

print(combined_table)

```
As we have learned from @tbl-modelresults and which we can see in @sec-index it is obvious that the MVP is not always the player that performs the best statistically, as LeBron James would have won 8 MVP awards before entering the 2013-2014 season, which would be 8 MVP awards in 10 years. Therefore, there must other factors excluding the statistics we were looking at, such as human error or prejudice, team success, and overall story telling which we will look further into in @sec-discussion.

# Results {#sec-results}

In this section, we examine key basketball statistics since 1986, beginning with the MVP Index, followed by points per game (PPG), rebounds per game (RPG), and assists per game (APG). The graphs will compare historical MVPs with players from each year who had the highest MVP Index. Analyzing these trends will provide insight into how these statistics have evolved over time among the NBA's elite players and may reveal which statistics are valued more in practice compared to the official MVP selection process.

## MVP Index Since 1986 {#sec-mvp-index}

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-mvpindex
#| fig-cap: "MVP index of of Historical and simulated MVPs"

merged_data <- bind_rows(
  merged_real_mvp %>% mutate(Source = "Historical MVPs"),
  mvp_table %>% mutate(Source = "Simulated MVPs")
)

# Create the faceted plot with lines connecting points
ggplot(merged_data, aes(x = Year, y = MVP_index, group = Source)) +
  geom_line(aes(color = Source), size = 1, alpha = 0.7) +  # Add lines connecting points
  geom_point(aes(color = Source), size = 3, alpha = 0.7) +  # Keep the points
  facet_wrap(~ Source, ncol = 2) +  # Facet by the Source column, side by side
  labs(
    title = "Scatter: Year vs. MVP Index",
    x = "Year",
    y = "MVP Index"
  ) +
  coord_cartesian(ylim = c(1.7, 2.8)) +  # Maintain the same y-axis range
  theme_minimal() +
  theme(
    legend.position = "none",  # Remove the legend since it's redundant with faceting
    strip.text = element_text(size = 12, face = "bold")
  )

```
In @fig-mvpindex, we observe that both graphs fluctuate significantly until the 2020s, where they converge around an MVP Index score of 2.2. This trend could suggest several possibilities. One explanation might be that MVPs are becoming less impactful, but this is highly unlikely and not supported by the evidence. A more plausible explanation is that the competition among players is becoming stronger and more consistent over time. It is important to remember that the MVP Index depends as much on the performance of all players in the dataset as it does on the individual player being evaluated.

For example, the 2004-05 and 2005-06 seasons show a notable dip in the MVP Index for historical MVPs. This can be attributed to Phoenix Suns' point guard Steve Nash, who won back-to-back MVP awards during those seasons. Interestingly, Nash did not rank in the top 10 for MVP Index, with two of his teammates, Amar'e Stoudemire and Shawn Marion, both achieving higher Index scores. While Nash's MVP selections remain some of the most debated in the past 20 years, it is important to consider the factors NBA analysts likely valued at the time. Nash's team experienced considerable success, even though they fell short of winning a championship. Furthermore, Nash was recognized for his exceptional leadership, serving as both a floor general and a mentor.

This raises critical questions: Did Nash truly deserve those MVP awards? And, perhaps more importantly, is the selection process for determining the MVP inherently flawed?

## PPG since 1986

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-pts
#| fig-cap: "MVP PPG over Time"

# Add a source column to the datasets
mvp_table <- mvp_table %>%
  mutate(Source = "Simulated MVPs")

merged_real_mvp <- merged_real_mvp %>%
  mutate(Source = "Historical MVPs")

# Combine datasets
combined_data <- bind_rows(
  mvp_table %>% select(Year, PTS, TRB, AST, Source),
  merged_real_mvp %>% select(Year, PTS, TRB, AST, Source)
)

# Plot for PTS
pts_plot <- ggplot(combined_data, aes(x = Year, y = PTS, group = Source, color = Source)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ Source) +
  labs(
    x = "Year",
    y = "PTS"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Plot for TRB
trb_plot <- ggplot(combined_data, aes(x = Year, y = TRB, group = Source, color = Source)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ Source) +
  labs(
    x = "Year",
    y = "TRB"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Plot for AST
ast_plot <- ggplot(combined_data, aes(x = Year, y = AST, group = Source, color = Source)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ Source) +
  labs(
    x = "Year",
    y = "AST"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Display plots separately
print(pts_plot)

```

We will now examine @fig-pts, which illustrates the points per game (PPG) of MVPs over the years. Similar to @fig-mvpindex, this graph shows fluctuations in the data; however, unlike the MVP Index, the values do not appear to converge around a single point. Instead, this graph reveals an interesting trend: for both Simulated MVPs and Historical MVPs, there seems to be an increasing baseline as well as an increasing upper limit. While the data points fluctuate year to year, the overall range of PPG among MVPs is expanding, with both higher peaks and higher troughs.

A notable exception to this trend is Steve Nash, the only MVP since 1986 to win the award while scoring under 20 PPG, as discussed in @sec-mvp-index. Excluding this outlier, the graphs suggest that over time MVPs, whether simulated or historical, are scoring more points on average, even those players who traditionally score less. This observation aligns with findings from [@yan], confirming that scoring among top-tier players has been steadily increasing over the years.

## RPG since 1986

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-reb
#| fig-cap: "MVP RPG over Time"

print(trb_plot)
```
In @fig-reb, we examine rebounds per game, providing additional insight into the type of MVPs and what was prioritized at different points in time. There is a notable dip in rebounds between the mid-2000s and the mid-2010s in both the Historical Data and Simulated MVPs. This period reflects a lack of "big men," such as centers and power forwards, who traditionally dominate rebounds on their teams. Until the late 2010s, rebounding was largely the domain of these larger players, but during this time, there seems to have been a vacuum of elite big men.

The early 2000s saw dominant rebounders like Shaquille O’Neal, Tim Duncan, and Kevin Garnett. However, their era was followed by a noticeable void until the emergence of smaller players, such as Russell Westbrook, who began recording higher rebound numbers. In the late 2010s and early 2020s, elite big men like Giannis Antetokounmpo, Nikola Jokic, and Joel Embiid rose to prominence. One of these three players has won the MVP award in each of the last six years, which accounts for the significant increase in rebounds observed in the later years of the graph.

## APG since 1986

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-ast
#| fig-cap: "MVP ASG over Time"

print(ast_plot)
```

Lastly, we examine @fig-ast, which graphs assists per game over time since 1986. Similar to @fig-reb, there is a dip in assists per game among both Historical MVPs and Simulated MVPs during the 1990s and early 2000s. This period was dominated by big men, who were known more for their rebounding abilities than their passing skills. However, we see a gradual increase in assists per game during the 2010s and 2020s, with a small outlier in Joel Embiid in 2023.

After reviewing these visuals, we can conclude that each of these statistics exhibits a somewhat growing trend. While the growth may not always be consistent, it does appear to exist. This suggests that MVPs in the modern NBA are increasingly likely to be selected if they are well-rounded players, athletes who not only score points and rebound effectively but also contribute as unselfish teammates by creating scoring opportunities through their passing.

# Discussion {#sec-discussion}

## Human Error and Bias {#sec-bias}

We must remember that, at the end of the day, the MVP is selected by a group of NBA analysts who come together once a year to cast their ballots for the player they believe deserves the award. It is also important to note that the award is called the "Most Valuable Player" and not the "Best Player" award, emphasizing that the recipient is deemed the most valuable to their team, rather than simply the best overall player. While the human element in voting allows analysts to consider both quantifiable statistics and intangible factors, such as leadership, teammate morale, and on- and off-court behavior, this subjective process introduces room for bias. Machines, in contrast, would likely base their decisions on a purely quantitative measure, such as an MVP Index similar to ours, and choose the player with the highest score.

Human bias can sometimes lead to controversial outcomes. For example, in 2016, Stephen Curry became the first unanimous MVP, receiving all 131 votes. In contrast, in 2013, LeBron James received 120 out of 121 votes, with the remaining vote going to Carmelo Anthony. While Carmelo had a strong season, LeBron’s performance that year was exceptional, combining statistical dominance with a second NBA championship win against one of the most well-constructed teams in league history. This example underscores that even a trustworthy and well intentioned group of analysts is still human and prone to occasional error or bias.

Additionally, the selection of an MVP can sometimes reflect narrative-driven decisions that align with storytelling or commercial interests. For instance, when Derrick Rose won the MVP in 2011, few would argue he was the best player in the world. By nearly every statistical measure, LeBron James had a superior season, and his team, the Miami Heat, achieved greater playoff success. However, Rose’s story resonated with fans—he was the hometown hero of Chicago, a city looking to rekindle its basketball glory from the Michael Jordan era. This compelling narrative undoubtedly contributed to his MVP selection, illustrating how storytelling can occasionally influence outcomes in ways that deviate from purely objective criteria.

## Consistency vs. Peak Performance {#sec-peak}

Averages do not show everything. It is possible for a player to be incredibly inconsistent for their team, playing well against weaker opponents while struggling against stronger ones. This could lead to impressive average stats that fail to reflect their performance when the team needed them most. On the other hand, peak performance, where a player dominates in key moments, often captures attention and can overshadow an otherwise inconsistent season. The MVP debate often centers around this tension between reliability and brilliance.A player who delivers steady, high-level contributions throughout the season provides a reliable foundation for their team’s success. These players might not always have headline-grabbing moments, but their ability to perform well in nearly every game can be the difference between making the playoffs or falling short. Their contributions build trust among teammates and coaches, as they can be relied upon to deliver in critical situations time and time again.

In contrast, players who showcase moments of sheer brilliance often become the face of the league. These athletes might not exhibit the same game to game reliability, but their ability to dominate in spectacular fashion leaves a lasting impression on fans and voters alike. Peak performers can change the momentum of a season with game-winning shots, record-breaking performances, or memorable battles against top opponents. However, their inconsistency can leave teams vulnerable during less-publicized matchups or critical stretches of the season.Ultimately, the MVP conversation remains an evolving debate, shaped by both quantitative measures and subjective interpretations of what makes a player "most valuable."

## Voting Methodology and Transparency {#sec-method}

The voting methodology and transparency in the NBA, particularly for awards like the MVP, have significant implications for the league’s credibility and fan engagement. The MVP award is determined by a panel of sportswriters and broadcasters from the United States and Canada, with each voter ranking their top five candidates. Points are assigned on a weighted scale, and the player with the highest total is named MVP. While this system allows for expert analysis, transparency remains a recurring discussion point. Critics often call for more public access to the detailed ballots to ensure accountability and minimize potential biases. The NBA has made strides in recent years by releasing voters’ choices post-award, fostering a culture of openness. However, debates about the inclusion of diverse perspectives, potential regional biases, and the weight given to narrative versus statistical performance continue to shape the conversation around fairness and transparency in the voting process.

## Weaknesses and next steps {#sec-weak}

First, what would have been ideal especially in terms of using our model, would have been to take players from the all 3 All NBA teams, and download the data from these players and have our model analyze the data so we would have an actual name of a player for our predicted MVP Index and not just the score. While this would have been great time constraints and other external factors hindered us from implementing this further analysis.

Second, for a more accurate MVP Index we could have sampled the entire league instead of the top 20-30 players, as this would have made a more accurate model as well as made the great players look much more exceptional with regards to the MVP Index, as we have only sampled players who have received at least one vote for an All NBA team, this leads to the average statistics we used to normalize our stats. This ended up making the Phenomenal players look great, and the great players look average, and the good players look sub par.

Next, team records are also incredibly important for MVP, if a player played incredibly one season but barely made the playoffs, or even did not at all, the player likely should not be receiving the MVP award, as how much worse could said team get if they weren't that good with the so called "best" player in the league. Unfortunately we were notable to inclue team record, team win percentage, or even player win percentage if we do not want to include the games a player may have missed. Unfortunately this data was not included in either type of data set that we sourced from [@bbref], and due to time constraints downloading potentially dozens if not hundreds additional data sets did not seem to be the best use of time in our case.

Lastly, as mentioned in @sec-bias the MVP award does not solely rely on statistics and quantifiable factors. While, stats are heavily used in the decision for choosing the MVP there are other. non tangible factors that must be taken into account, other wise incredibly selfish players who do not help their teams have a much higher chance of winning.

Our next steps, would like be to monitor this NBA season closely potentially add some of the data mentioned earlier in @sec-weak, such as win percentage, and individual historical statistics to more accurate predict the model. We would also like to revisit this paper tat the end of this NBA season once the NBA awards are announced and see how we could improve our model and how accurate it was in predicting the MVP Index of the next MVP.

\newpage

\appendix

# Appendix {.unnumbered}

# Surveys, Sampling, and Observational Data

## Overview of MVP Voting Methodology

The MVP award voting process relies on a panel of sportswriters and broadcasters across the United States and Canada, with each voter ranking their top five players. The point system allocates 10 points for a first-place vote, 7 for second, 5 for third, 3 for fourth, and 1 for fifth. The player with the highest total points is named the MVP. While this system ensures broad coverage across media markets, it introduces potential biases rooted in the selection of voters, the weighting system, and the lack of transparency in the criteria used by individual voters [@schmitz].

## Observational Nature of NBA Data

MVP voting data is inherently observational, reflecting subjective decisions rather than controlled experiments. This reliance on subjective assessments introduces variability, as voters may prioritize narrative driven criteria such as team success, leadership, or memorable moments over statistical performance. Observational data like this cannot account for unmeasured factors, such as voters’ implicit biases or media-driven narratives. Additionally, players on historically successful teams or those in larger markets may receive disproportionate attention, skewing the outcomes.

## Sampling Representation of Voters

The panel of voters is not a random sample of all stakeholders in the NBA community. Instead, it is a curated group of media professionals, which may underrepresent certain perspectives, such as fans, coaches, or analytics experts. This introduces a potential sampling bias that could influence the results. For example, regional biases might favor players from large-market teams, while recency bias could elevate players with late-season surges. Simulation studies could explore the extent of these biases by randomly reweighting votes or introducing new voter segments, e.g., analytics specialists.

##Simulation of MVP Voting Under Different Scenarios

To examine the robustness of MVP outcomes, simulations were conducted to evaluate how changes to the voting methodology could influence results. For instance, adjusting the weight given to statistical performance in vote tallies revealed several seasons where the MVP Index-aligned winner differed from the historically selected MVP. Similarly, experimenting with alternative weighting schemes, such as assigning equal points for all top-five rankings instead of a tiered system, led to notable shifts in the final outcomes. These simulations highlight the significant role that methodological choices play in shaping MVP selections, showing that even small adjustments can result in different winners. By simulating different voting scenarios, it becomes clear that the current system is sensitive to how votes are counted and weighted, underscoring the need for transparency and careful consideration of methodology.

# Table By MVP Index {#sec-index}

```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false

print(mvp_table)
```

# Historical MVPs

```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false

Real_MVP_Datas <- Real_MVP_Data[-c(1, nrow(Real_MVP_Data)), ]

write.csv(Real_MVP_Datas, "../data/02-analysis_data/Real_MVP_Datas.csv", row.names = FALSE)

print(Real_MVP_Datas)
```

# Model details {#sec-model-details}

```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false
lm_summary <- summary(lm_model)

print(lm_summary)
```

\newpage

# References